Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-06-27 code/rmarkdown/article.Rmd
+ progress on the discussion + also included some stats in the discussion

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environment variation it also secured was found to depend on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56). The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($> 80$ %). Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). Compared to the environmental variables, the slopes of these relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($> 95$ %). Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of their intra-specific genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_). These results suggest while prioritizations generated using environmental and geographic targets tend to secure more genetic variation, prioritizations generated using only these targets can sometimes secure only a small fraction of intra-specific genetic variation.

@@

## Discussion
* introduction
We aimed  to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning. Our study is the first to address this hypothesis using moderately realistic planning scenarios. Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species. Furthermore, we found that the addition of environmental and geographic targets to the planning process resulted in prioritizations that generally secured a greater proportion of adaptive and neutral genetic variation.

* consequences for conservation planning

* outlier detection methods

* mapping genetic variation

* limitations/future directions
We wish to alert our readers to several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2). We used this resolution because the genetic data were collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have artificially introduced additional spatial auto-correlation into the dataset. Second, we used geographic distances as surrogates for connectivity between planning units. Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg. @r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters. Third, we used AFLP data to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation in order to generate prioritizations in a feasible amount of time. Overall, we are confident that the limitations associated with our study are relatively minor.

* Concluding remarks
The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning. We found that environmental and geographic are generally effective surrogates for most species. Although prioritizations generated using amount-only targets tended to secure most of the adaptive variation for most species, the addition of surrogate targets to the planning process substantially improved the proportion of intra-specific neutral genetic variation. Conservation planners are therefore urged to include these surrogates in conservation planning exercises.

>>>>

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environment variation it also secured was found to depend on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). After establishing  a relationship between the proportion of the surrogates secured in a prioritisation and proportion of genetic variation secured, the next step was to determine if prioritisations generated under various scenarios benefitted from the addition of surrogate-based targets.

@@

## Discussion
<!--  introduction -->
We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning. Our study is the first to address this using a large-scale data set. Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species. Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritizations that secured a greater proportion of adaptive and neutral genetic variation for most species.

<!-- surrogate/genetic dissimilarity matrix correlations -->
Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1). For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs. adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` $\pm$ SD; geographic vs. neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` $\pm$ SD) were generally much higher than the marginal $R^2$ values (environmental vs. adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` $\pm$ SD; geographic vs. neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` $\pm$ SD). This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates. In particular, the large differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic distances suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals. This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX]. After verifying that the spatial pattern of variation in the surrogates generally correlated with patterns of intra-specific adaptive and neutral genetic variation, our next step was to determine if this pattern of correlation was strong enough to operationalise for conservation planning.
Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1). For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs. adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` $\pm$ SD; geographic vs. neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` $\pm$ SD) were generally much higher than the marginal $R^2$ values (environmental vs. adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` $\pm$ SD; geographic vs. neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` $\pm$ SD). This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates. In particular, the large differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic distances suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals. This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX]. After verifying that the spatial pattern of variation in the surrogates generally correlated with patterns of intra-specific adaptive and neutral genetic variation, our next step was to determine if this pattern of correlation was strong enough to operationalise for conservation planning.

<!-- surrogate/genetic held correlations -->
Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2). Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg. _Gypsophila repens_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species. The mechanism for this is two fold. In such alpine landscapes such as those investigated in this study, connectivity between areas can be strongly affected by local topology [@XXXX]. Although  topographic-based variables describing the connectivity between areas may offer a better surrogate (eg. least cost paths based on elevation; @rXXX), such variables need to scaled to accommodate non-linear relationships between the variable and connectivity.  and the appropriate scalings vary immensely between species--especially between plant species which can make use of remarkably different dispersal strategies (eg. wind dispersal, animal dispersal, offshoots from the mother plant). As a consequence, more complex surrogates of connectivity generally need to be tuned for each species--which requires either an in-depth understnading of the study species [eg. @XXX] and/or genetic data [eg. @XXX]--and so are not particularly applicable for broad-scale prioritisations that involve hundreds of species.
Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2). Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg. _Gypsophila repens_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species. The mechanism for this is two-fold. First, the plant species investigated in this study utilise a diverse set of dispersal mechanisms, including anemochorous (eg. wind-based dispersal), zoochorous (animal-based dispersal), and boleochorous (gravity-mediated dispersal). Plant species utilising anemochorous and zoochorous strategies can exhibit remarkable long-distance dispersal capabilities [@rXXX], and geographic distance-based surrogates would fail to describe this. Second, in alpine landscapes such as those investigated in this study, connectivity between areas can be strongly affected by local topology [@XXXX]. However, distance measures based on landscape features (eg. elevation) require non-linear scaling--based on an in-depth understanding of the study species [eg. @XXX] and/or genetic data [eg. @XXX]--to be effective surrogates for connectivity, and so are not applicable for large-scale prioritisations that involve hundreds of species. Therefore, we conclude that the surrogates generally performed well enough considering their lack of focus to any particular species.

<!-- suitable targets for surrogates -->
Generally, the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57). To secure a representative sample of each species' adaptive genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` % $\pm$ `r sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` SD of the environmental conditions found across the planning units occupied by the species. On the other hand, to secure a representative sample of the species' neutral genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` $\pm$ `r sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` SD of the variation in geographic position among the planning units occupied by the species. These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation. Additionally, these results suggest using surrogate-based targets alone to generate prioritisations may be insufficient to secure a representative sample of species' genetic variation. For all species, prioritisations were generated that secured a high proportion of the surrogate variable (> 90 %) but also secured a minuscule proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57). Thus we recommend that prioritisations be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets. After establishing a relationship between the proportion of surrogate variation and genetic variation secured in prioritisation, we than generated prioritisations under various planning scenarios to provide insight into how the surrogates may work in the real-world.

<!-- adaptive genetic variation secured in prioritisations-->
Almost all prioritisations secured a representative sample of species' adaptive genetic variation--regardless of whether surrogate-based or genetic targets were used or not (Figure 4). This result suggests that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount targets. The spatial patterns of adaptive genetic variation tended to cluster into a few distinct groups for most species (eg. XXX; XXX; XXX; XXX), and so a prioritisation would only need to site a single planning unit in each cluster to secure a representative sample of the species' adaptive genetic variation. As a consequence, the single-species prioritisations generated using amount targets and without consideration to opportunity cost were often able to achieve high representation due to the fact they were essentially random selection (Figure 4a). To a lesser extent this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs are fairly homogenous among planning units with exception to a few planning units lowland areas.

Generally, all the multi-species prioritisations secured a representative sample of the species' intra-specific adaptive genetic variation (Figures 4e and 4g). Unlike the single-species prioritisations, the multi-species prioiritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes. Since these prioritisations were forced to secure at least `r rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100` % of land occupied each species, these prioritisations were forced to secured planning units a range of environmental conditions due to the presence of habitat specialists in the data set (eg. _Cerastium uniflorum_ and _Androsace obtusifolia_). Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used. This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning [@rXXX].

<!-- neutral genetic variation secured in prioritisations-->
Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4). With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h). Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species. Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation. Furthermore, this improvement was associated with only a minor increase in cost (XXXX)--emphasizing the utility of surrogates in conservation.

<!--  limitations/ weaknesses -->
We wish to alert our readers to several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2). We used this resolution because the genetic data were collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have artificially introduced additional spatial auto-correlation into the data set. Second, we used geographic distances as surrogates for connectivity between planning units. But as previously, using more geographic distances were found to be adequate and more complex measures of connectivity cannot generally be used in more large-scale planning scenarios. Third, we used AFLP data to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation in order to generate prioritizations in a feasible amount of time. Overall, we are confident that the limitations associated with our study are relatively minor.

<!--  conclusion -->
The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning. We found that environmental and geographic are generally effective surrogates for most species. Although prioritizations generated using amount-only targets tended to secure most of the adaptive variation for most species, the addition of surrogate targets to the planning process substantially improved the proportion of intra-specific neutral genetic variation. Conservation planners are therefore urged to include these surrogates in conservation planning exercises.
